Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance

### 2023-05-06 Manuscript/Manuscript.Rmd
MS: edits to text, reordering NIT-GC sections - tc

The ability to perceive changes in the environment and adapt behaviour accordingly is essential for the survival of all living organisms. Avoidance from UV radiation, which damages DNA essential for survival, is proposed to be one of the behaviours that existed before the abundant development of Earth's oxygen to block UV radiation in the atmosphere (Pittendrigh). In the ocean, almost all zooplankton have a pronounced rhythmic behaviour called Diel vertical migration (DVM), which consists of upwards towards the surface at dusk and descend back to deeper water before dawn. Avoidance behaviour to escape harmful UV radiation during the day is closely linked to the control of vertical migration in the ocean. However, little is known about the molecular neural circuit mechanisms in UV avoidance behaviour in many animals, including zooplankton. In neural circuits in general light response, the activation of various receptors by excitatory and inhibitory neurotransmitters such as glutamate, GABA and dopamine is known in various species. In addition to this, it has long been known that in nitric oxide (NO), as a neuromodulatory molecule, it is also important as a modulator of visual activity. NO is a small free-radical molecule with diffusible and changeable characteristics, and thus works as one of the neural signalling molecules. The function of NO in the neural circuitry of vision has been well studied in vertebrates. In the mammalian retina, NO is an essential signalling molecule in all retinal cells. NO produced by nitric oxide synthase (NOS) in amacrine cells, one of the retinal cells, activates *soluble guanylate cyclase (sGC)* expressed in cone bipolar cells, which produces cGMP and regulates downstream signalling. We therefore hypothesise that NO-mediated neural signalling may be involved in the photoreceptor circuits that control UV avoidance behaviour. To test this hypothesis we used the annelid *Platynereis dumerilii* **(Figure 1A)**, which conserved vertebrate-type ciliary photoreceptors (cPRCs) [@Arendt2004]. The cPRC expresses *c-opsin1*, which absorbs UV radiation, is involved in melatonin production and regulates circadian swimming behaviour [@Tosches2014; @tsukamoto2017]. Furthermore, we previously reported abnormal UV avoidance behaviour in *c-opsin1* knockout larvae [@Veraszto2018]. The cPRCs express the UV-absorbing c-opsin1 and mediate UV avoidance in the larvae [@Veraszto2018]. In adult worms, UV light and c-opsin1 modulate locomotion and maturation [@Veedin2021]. These suggest that Platynereis larvae are a highly suitable model for elucidating the neural circuit mechanisms of NO-induced UV avoidance behaviour, and the accessibility of genome editing, behavioural assays, calcium imaging and whole-body connectomics could be very useful for study at the molecular level[@ozpolat2021].

@@

We identified a single *nitric oxide synthase* (*NOS*) gene in the *Platynereis dumerilii* genome and transcriptome data. Phylogenetic analysis of NOS proteins indicate that *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences **(Figure 1---figure supplement 1)**. To characterise the expression pattern of *NOS* we used in situ hybridization chain reaction (HCR) and transient transgenesis. In two- and three-day-old larvae, we detected *NOS* expression in four cells (two of them weakly expressing) in the apical organ region **(Figure 1D and Figure 1---figure supplement 2)**. The same four cells were labelled with a reporter construct driving palmitoylated-tdTomato by a 12 kbp upstream genomic region of the *NOS* gene **(Figure 1E)**. This reporter also revealed the axonal projections of the NOS-expressing neurons. The position and morphology of the four *NOS*-expressing cells allowed us to identify the same cells in the three-day-old whole-body volume EM data [@williams2017; @veraszto2020] as four interneurons (INNOS) **(Figure 1B,C)**. The INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGW interneurons, which are also cPRC targets **(Figure 1C,F)**. INRGW neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1), which synapse on the prototroch ciliary band and the cholinergic MC ciliomotor neurons **(Figure 1C,F)**. We also detected *NOS* expression by HCR in the region of the visual eyes (adult eyes) and the pigmented eyespots **(Figure 1---figure supplement 1)**.

@@

The specific expression of *NOS* in interneurons directly postsynaptic to the cPRCs suggests that NO signalling is involved in the UV response. To test if NO is produced during UV/violet responses, we injected zygotes with the fluorescent NO-reporter DAF-FM and imaged two-day-old larvae during the local stimulation of the cPRC cilia by 405 nm light. Following light stimulation, we detected an increase in DAF-FM fluorescence in the region of the INNOS cells **(Figure 2)**.

@@

To test this, we generated two *Platynereis* *NOS* knockout lines with the CRISPR/Cas9 method. We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both leading to an early stop codon and thus likely representing null alleles **(Figure 3A)**. We could establish a homozygous line for both mutations indicating that NOS is not an essential gene. To quantify UV avoidance, we recorded the trajectories of freely swimming larvae in vertical columns, illuminated laterally from two opposite sides with 395 nm UV/violet light. To analyse the behaviour of larvae before and after UV stimulation, the temporal changes in some parameters were compared using tracking of individual larvae of the WT and mutant **(Figure 3B,C and Figure 3---figure supplement 1A-D)**. The results showed that in the WT there was a rapid and strong increase in downward swimming approximately 15 s after UV stimulation, whereas this did not occur in the mutant **(Figure 3C and Figure 3---figure supplement 1B)**. Interestingly, there was already a difference in swimming speed between the WT and the mutant before UV stimulation (in the dark) **(Figure 3---figure supplement 1C,D)**. After UV stimulation, the swimming speed of the WT increased and reached a maximum after about 20 s. In contrast, the mutants showed a slight increase in speed after UV stimulation, but swimming speed was not maintained and returned to the speed before UV stimulation. We also tested phototaxis, by exposing larvae to 480 nm directional and collimated light from the top of the column. As previously shown, wild-type larvae swim downward within 30 sec following non-directional UV/violet light stimulation [@Veraszto2018]. In contrast, homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response **(Figure 3D and Figure 3---figure supplement 1E)**. This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae and reveals a requirement for NOS in UV-avoidance behaviour [@Veraszto2018]. *NOS*-mutant larvae also showed reduced phototactic behaviour, suggesting a function in the visual eyes mediating phototaxis [@Randel2014a], where *NOS* is also expressed **(Figure 1---figure supplement 1A,B)**. To test if the observed defects are due to a role of NOS in neuronal signalling or circuit developmental, we also tested larvae exposed to a NOS inhibitor, L-NAME. Larvae incubated for 5 min in 0.1 mM or 1 mM L-NAME showed defective UV avoidance, but normal phototaxis **(Figure 3E)**. Overall, our results indicate an acute requirement for NOS signalling in UV-avoidance and a possible indirect, developmental role in the visual system. To investigate the role of NO on UV avoidance behaviour, receptors for NO were explored.

@@

@@

NO generally acts via soluble guanylate cyclases (sGC), belonging to the guanylate cyclase family with a CYC domain (PF00211). NO binding to the heme group of sGC leads to increased cyclic guanosine monophosphate (cGMP) production. Analysis of sGCs in *Platynereis* indicated that these genes are not expressed in any of the cells of the cPRC circuit [@Veraszto2017]. Recently, Moroz and coworkers reported an atypical but widely conserved family of guanylyl cyclases with a NIT (nitrite/nitrate sensing) domain (PF08376) (NIT-GC) as potential mediators of NO signalling [@Moroz2020]. To search for NIT-GCs in *Platynereis*, we searched transcriptome resources and retrieved 15 potential NIT-GC homologs **(Figure 4---figure supplement 1)**. To analyse the relationship of these sequences to metazoan NIT-GCs, we retrieved protein sequences with a CYC domain from the transcriptome and genome databases of 45 metazoan species and 2 choanoflagellate species. We carried out cluster analysis and did phylogenetic reconstruction on a group of membrane-bound guanylyl cyclases with sGCs as an outgroup. In agreement with Moroz et al. [@Moroz2020],we found a group of GCs with NIT domains with representatives in placozoans, cnidarians, some ecdysozoans, echinoderms, and lophotrochozoans. The 15 *Platynereis* sequences were parts of several major and ancient clades of the tree **(Figure 4---figure supplement 1)**. To characterise the expression of NIT-GCs, we using previously published spatially mapped single-cell transcriptome data [@Achim2015; @williams2017]. Among the 15 NIT-GCs, two showed high and specific expression in the cPRCs and one was expressed in the INNOS cells **(Figure 4---figure supplement 1)**. In the single-cell data, we could identify the cPRCs by the specific expression of c-opsin1 and a pedal-peptide2 neuropeptide precursor (MLD proneuropeptide) [@Arendt2004; @williams2017], which have previously been described as cPRC markers **(Figure 4---figure supplement 2A)**. The INNOS cells were identified by NOS expression and their spatial mapping. We decided to focus on two NIT-GCs expressed in the cPRCs and with a full-length sequences, NIT-GC1 and NIT-GC2. To confirm the single-cell data, we first carried out in situ HCR with probes for *NIT-GC1* and *NIT-GC2* mRNA. Both genes showed very clear and specific expression in the four cPRCs, as confirmed by co-labeling with an acetylated α-tubulin antibody and with an HCR probe against *MLD proneuropeptide* **(Figure 4A,B and Figure 4---figure supplement 2A-C)**. To analyse the subcellular localisation of NIT-GC1 and NIT-GC2 at the protein level, we raised and affinity-purified polyclonal antibodies against a specific peptide sequence from both proteins. In immunostainings, we found that NIT-GC1 was localise to the region corresponding to the axonal projections of the cPRCs in the anterior nerve plexus **(Figure 4C)**. Co-immunostaining with the rabbit NIT-GC1 and the rat NOS antibodies revealed that both proteins were localised in close proximity in the neurosecretory plexus **(Figure 4---figure supplement 2D)**. In contrast, NIT-GC2 strongly and specifically labelled the large sensory ciliary region of the cPRCs **(Figure 4D)**. These different subcellular localisations suggest that the two NIT-GCs are involved in different intracellular signalling processes in the ciliary and axonal regions of the cPRCs.

**NIT-GC1 is required for the depolarising response of cPRC to UV stimulation**

To investigate how NIT-GCs-mediated NO signalling is involved in cPRC activation, the calcium sensor GCaMP6s was ubiquitously expressed in larvae and calcium imaging in the cPRC during UV/violet stimulation. As we have shown previously, a 20-sec local stimulation of the ciliated part of the cPRC with 405 nm light lead to a transient decrease in cPRC calcium levels, likely due to the hyperpolarisation of the cPRCs [@Veraszto2018]. After 20-sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus - a response that may involve depolarisation. This activation phase occurs after the end of the stimulation and may be due to a slower neuroendocrine feedback [@Veraszto2018]. First, to determine whether NO mediates such a feedback, we measured calcium levels in *NOS*-mutant larvae. While we detected the inhibition phase in cPRC in *NOS*-mutants, this was not followed by activation and calcium levels remained low **(Figure 4E,F)**. These results revealed an essential requirement for NO signalling in the second phase of the cPRC response. Next, to characterise the role of NIT-GCs in changes of calcium levels in cPRC, calcium imaging was performed when each *NIT-GC* gene was knocked down using morpholino **(Figure 4---figure supplement 2E,F)**. In NIT-GC1 morphant larvae, 405 nm light stimulation caused a decrease in calcium levels in the cPRC, but the increase in calcium levels that occurred after the light stimulation was stopped was strongly suppressed **(Figure 4G)**. This result is similar to the response of NOS mutants in cPRC, suggesting that NIT-GC1 mediates depolarising responses after UV stimulation of cPRC by NO and may be involved in extracellular signalling. Interestingly, in NIT-GC2 morphant larvae, an increase in calcium levels occurred immediately after light stimulation **(Figure 4H)**. The different responses of cPRC upon knockdown of NIT-GC1 and NIT-GC2 suggest that there may be different intracellular signalling mechanisms associated with their respective localisations. We focused our analysis on NIT-GC1 because it was localised to cPRC neuroaxons within the neuroplexas region and is involved in extracellular signalling after UV stimulation. It is generally known that depolarisation in the PRC involves elevated cGMPs in the photoreceptors. We therefore used an in vitro system in which cultured cells expressed NIT-GC1 to assess the cGMP activity of NIT-GC1.

@@

**NIT-GC1 produces cGMP by nitric oxide**

NIT-GC1 has a highly conserved cyclase domain and activation of NIT-GC1 is expected to produce cGMP **(Figure 4I)**. Recently, a synthetic protein called Green cGull was developed as a cGMP indicator that can be used in cultured cell systems [@Matsuda2017]. We co-expressed Green cGull and NIT-GC1 using the 2A self-cleaving peptide in cultured cells COS-7 (monkey kidney) and tested the intracellular dynamics of cGMP in the activation of NIT-GC1 **(Figure 4J)**. When the NO donor SNAP was applied to the NIT-GC1 and Green cGull co-expression cells, the intracellular cGMP levels was significantly increased compared to when NIT-GC1 was not expressed **(Figure 4K,L)**. The increase in fluorescence intensity also did not occur when DMSO was applied **(Figure 4M)**. To confirm whether this increase in intracellular cGMP levels is NIT domain-dependent response of NIT-GC1, mutant form of NIT-GC1 without the NIT domain was expressed and SNAP was applied **(Figure 4I)**. As a result, the increase in fluorescence intensity was significantly reduced **(Figure 4N)**. These results indicate that NIT-GC1 produces cGMP via the NIT domain. In bacteria, it has been proposed that the NIT domain regulates cellular functions in response to changes in extracellular and intracellular nitrate and nitrite concentrations. Our results suggest that NIT-GC1 may be sensitive to NO. This could be a possible feedback pathway in which NIT-GC1 expressed in cPRC is activated by NO produced by INNOS. In order to further explore the role of NO in the cPRC circuit, we therefore compared calcium imaging of postsynaptic cells (INNOS, INRGW, Ser-h1 and MC) in the cPRC circuit during UV stimulation of cPRC between WT and NOS mutant larvae.

@@

>>>>

Neuronally produced nitric oxide (NO) can act as a diffusible neuromodulator. This small free-radical molecule is synthesized by nitric oxide synthase (NOS)...

In the mammalian retina, NO is produced by amacrine cells and activates a *soluble guanylate cyclase (sGC)* in cone bipolar cells. The NO-dependent increased in cGMP tunes bipolar cell activity .... [refs]
[other examples]

To investigate NO signalling in the modulation of neural circuit dynamics, we studied the larval stages of the marine annelid model *Platynereis dumerilii* [@ozpolat2021]. We uncovered an essential function for NO in these larvae in UV/violet-light avoidance.
In *Platynereis* larvae, UV/violet avoidance is characterised by downward swimming and is mediated by the UV/violet-responding brain ciliary photoreceptor cells (cPRCs).

The cPRCs express a vertebrate-type ciliary opsin, c-opsin1 [@Arendt2004]. This opsin forms a UV-absorbing bistable photopigment with an absorption maximum around 384 nm [@tsukamoto2017, @Veraszto2018]. *c-opsin1* knockout larvae show defective UV/violet avoidance [@Veraszto2018]. 
Here we show that *Platynereis* *NOS* is expressed in interneurons of the cPRC circuitry and is required for UV/violet-avoidance behaviour. By combining calcium imaging across the fully-mapped cPRC circuit, genetic perturbations and mathematical modelling, we describe how NO tunes circuit dynamics through non-synaptic reverse signalling to cPRCs. This delayed neuroendocrine feedback is essential for circuit output, and integrates UV/violet input and also induces a short-term memory manifest in altered circuit activity.

#In adult worms, UV light and c-opsin1 modulate locomotion and maturation [@Veedin2021].

@@

We identified a single *nitric oxide synthase* (*NOS*) gene in the *Platynereis dumerilii* genome and transcriptome data. Phylogenetic analysis of NOS proteins indicate that *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences **(Figure 1---figure supplement 1)**. To characterise the expression pattern of *NOS* we used in situ hybridization chain reaction (HCR) and transient transgenesis. In two- and three-day-old larvae, we detected *NOS* expression in four cells (two of them weakly expressing) in the apical organ region **(Figure 1D and Figure 1---figure supplement 2)**. *NOS* was also expressed in the region of the visual eyes (adult eyes) and the pigmented eyespots **(Figure 1---figure supplement 1)**. The four apical organ cells, but not the eyes, were also labelled with a *NOS* reporter construct driving palmitoylated-tdTomato **(Figure 1E)**. This reporter also revealed the axonal projections of these central NOS-expressing neurons. The position and morphology of the four *NOS+* cells allowed us to identify the same four cells as four interneurons (INNOS) in our three-day-old whole-body *Platynereis* volume EM data [@williams2017; @veraszto2020] **(Figure 1B,C)**. In the synaptic connectome, the INNOS cells are postsynaptic to the UV-sensory cPRCs and presynaptic to the INRGW interneurons, which are also cPRC targets **(Figure 1C,F)**. INRGW neurons synapse on the head serotonergic ciliomotor neurons (Ser-h1), which synapse on the prototroch ciliary band and the cholinergic MC ciliomotor neurons **(Figure 1C,F)**.

@@

The expression of *NOS* in the INNOS interneurons in the cPRC circuit suggests that NO signalling may be involved in UV/violet-avoidance. To test this, first we asked whether NO is produced during UV/violet stimulation of the larvae. We injected the fluorescent NO-reporter DAF-FM into zygotes and imaged two-day-old larvae while exposing the ramified sensory cilia of the cPRCs to 405 nm violet light. Following light stimulation of the cPRCs but not a control area, we detected an increase in DAF-FM fluorescence in the anterior neurosecretory neuropil, the region of INNOS projections **(Figure 2)**.

@@

Next, we wanted to directly test whether NO signalling is involved in UV/violet avoidance. To achieve this, we generated two *Platynereis* *NOS* knockout lines with the CRISPR/Cas9 method. We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both leading to frame-shift and an early stop codon and thus likely representing null alleles **(Figure 3A)**. We could establish a homozygous line for both mutations indicating that NOS is not an essential gene. To quantify UV avoidance, we recorded the trajectories of freely swimming wild type and mutant larvae in vertical columns, illuminated laterally from two opposite sides with 395 nm UV light **(Figure 3B, C and Figure 3---figure supplement 1A-D)**. As previously shown, wild-type larvae swim downward following non-directional UV/violet light stimulation [@Veraszto2018]. In contrast, both two- and three-day-old homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response **(Figure 3B-D and Figure 3---figure supplement 1)**. This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae and reveals a requirement for NOS in UV-avoidance behaviour [@Veraszto2018]. Wild type but not mutant larvae also showed an increase in swimming speed under UV light that may be due to downward swimming trajectories (swimming with gravity) **(Figure 3---figure supplement 1C,D)**. 
We also tested directional phototaxis, by exposing larvae to 480 nm directional and collimated light from the top of the column. Three-day-old but not two-day-old *NOS*-mutant larvae also showed reduced phototactic behaviour, suggesting a function for *NOS* in the visual eyes that mediate three-day-old phototaxis [@Randel2014a] **(Figure 1---figure supplement 1A,B)**. 
To distinguish between an acute or developmental function of NOS in light responses, we next tested larvae exposed to the NOS inhibitor L-NAME. Larvae incubated for 5 min in 0.1 mM or 1 mM L-NAME showed a dose-dependent inhibition of UV avoidance. In contrast, phototaxis was not effected **(Figure 3E)**. Overall, our results indicate an acute requirement for NOS signalling in UV-avoidance and a possible indirect, developmental role in visual phototaxis.

@@

**NO feedback tunes cPRC responses to UV/violet stimulation**

To investigate how NO signalling alters the dynamics of the cPRC circuit, we carried out calcium imaging experiments. We ubiquitously expressed the calcium sensor GCaMP6s in larvae and imaged calcium signals during 405 nm stimulation of the cPRCs. As we have shown previously, a 20-sec local stimulation of cPRC cilia lead to a transient increase in cPRC calcium levels, followed by a transient decrease [@Veraszto2018]. After ~20-sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus -- a response that may involve depolarisation. This activation phase occurs after the 20 sec stimulation period and is likely due to a delayed neuroendocrine feedback [@Veraszto2018]. To determine whether NO mediates such a feedback, we repeated the experiment in *NOS*-mutant larvae. While we detected the initial activation phase followed by inhibition, in *NOS*-mutants this was not followed by delayed activation and calcium levels dropped to a low steady-state level **(Figure 4E,F)**. We thus identified a requirement for NO signalling in the late-phase activation of cPRCs.

@@

Next, we aimed to identify the NO-receptor in the cPRCs. NO generally acts via soluble guanylate cyclases (sGC), belonging to the guanylate cyclase family with a CYC domain (PFAM domain: PF00211). NO binding to the heme group of sGC leads to increased cyclic guanosine monophosphate (cGMP) production. Analysis of sGCs in *Platynereis* indicated that these genes are not expressed in any of the cells of the cPRC circuit [@Veraszto2017]. Recently, Moroz and coworkers reported an atypical but widely conserved family of guanylyl cyclases with a NIT (nitrite/nitrate sensing) domain (PF08376) (NIT-GC) as potential mediators of NO signalling [@Moroz2020]. To search for NIT-GCs in *Platynereis*, we searched transcriptome resources and retrieved 15 potential NIT-GC homologs **(Figure 4---figure supplement 1)**. To analyse the relationship of these sequences to metazoan NIT-GCs, we retrieved protein sequences with a CYC domain from the transcriptome and genome databases of 45 metazoan species and 2 choanoflagellate species. We carried out cluster analysis and did phylogenetic reconstruction on a group of membrane-bound guanylyl cyclases with sGCs as an outgroup. In agreement with Moroz et al. [@Moroz2020], we found a group of GCs with NIT domains with representatives in placozoans, cnidarians, some ecdysozoans, echinoderms, and lophotrochozoans. The 15 *Platynereis* sequences were parts of several major and ancient clades of the tree **(Figure 4---figure supplement 1)**. To characterise the expression of NIT-GCs, we used previously published spatially mapped single-cell transcriptome data [@Achim2015; @williams2017]. Among the 15 NIT-GCs, two showed high and specific expression in the cPRCs and one was expressed in the INNOS cells **(Figure 4---figure supplement 1)**. In the single-cell data, we could identify the cPRCs by the specific expression of c-opsin1 and a pedal-peptide2 neuropeptide precursor (MLD proneuropeptide) [@Arendt2004; @williams2017], which have previously been described as cPRC markers **(Figure 4---figure supplement 2A)**. The INNOS cells were identified by NOS expression and spatial mapping in the brain [@Achim2015]. We decided to focus on two NIT-GCs expressed in the cPRCs and with a full-length sequence, NIT-GC1 and NIT-GC2. To confirm the single-cell data, we first carried out in situ HCR with probes for *NIT-GC1* and *NIT-GC2* mRNA. Both genes were specifically expression in the four cPRCs, as confirmed by co-labeling with an acetylated α-tubulin antibody and with an HCR probe against *pedal peptide 2/MLD proneuropeptide* **(Figure 4A,B and Figure 4---figure supplement 2A-C)**. 
To analyse the subcellular localisation of NIT-GC1 and NIT-GC2 at the protein level, we raised and affinity-purified polyclonal antibodies against a specific peptide sequence from both proteins. In immunostainings, we found that NIT-GC1 was localise to the region corresponding to the axonal projections of the cPRCs in the anterior nerve plexus **(Figure 4C)**. Co-immunostaining with the rabbit NIT-GC1 antibody and a custom rat antibody raised against a fragment of *Platynereis* NOS, revealed that both proteins were localised in close proximity in the neurosecretory plexus **(Figure 4---figure supplement 2D)**. In contrast, NIT-GC2 specifically labelled the large sensory ciliary region of the cPRCs **(Figure 4D)**. These different subcellular localisations suggest that the two NIT-GCs are involved in different intracellular signalling processes in the ciliary and axonal regions of the cPRCs.

**NIT-GC1 produces cGMP in an NO-dependent manner**

To further characterise these atypical guanylyl cyclases, we focused on NIT-GC1 and carried out in vitro experiments. In bacteria, NIT domains are thought to regulate cellular functions in response to intra- or extracellular nitrate and nitrite. NIT-GC1 has a NIT domain and a highly conserved cyclase domain that is expected to catalyse cGMP synthesis **(Figure 4I)**. The NIT domain may render NIT-GC1 dependent on NO signals. To test this, we co-expressed the synthetic cGMP indicator Green cGull [@Matsuda2017] and NIT-GC1 in cultured cells COS-7 (monkey kidney). For balanced expression, we used a single plasmid with the two open-reading frames separated by the 2A self-cleaving peptide **(Figure 4J)**. Application of the NO donor SNAP lead to increased cGMP levels, an effect we did not observe when cells were exposed to DMSO or when Green cGull was expressed alone **(Figure 4K-M)**. To test whether cGMP production is dependent on the NIT domain, we also tested a deletion construct of NIT-GC1 lacking the NIT domain **(Figure 4I)**. Cells expressing this construct and Green cGull did not show increased cGMP levels when exposed to SNAP **(Figure 4N)**. These results indicate that NIT-GC1 is able to catalyse cGMP production in an NO-dependent manner and this function requires the NIT domain. These results establish NIT-GC1 as a biochemical NO sensor.

**NIT-GC1 is required for NO-mediated feedback to cPRCs during the UV response**

To test the in vivo function of NIT-GC1 and NIT-GC2 in cPRC responses, we returned to calcium imaging. 
To interfere with NIT-GC function, we used morpholino-mediated knockdowns. We used two translation-blocking morpholinos for each *NIT-GC* gene and tested knockdown efficiency by immunostaining injected animals with the NIT-GC1 and NIT-GC2 antibodies **(Figure 4---figure supplement 2E,F)**. For both genes, the morpholinos led to a strong reduction in the respective antibody signal, confirming that the morpholinos efficiently knock down translation and at the same time that the antibodies are specific.

In NIT-GC1 morphant larvae, 405 nm light stimulation caused a decrease in calcium levels in the cPRC, but the increase in calcium levels that occurred after the light stimulation was stopped was strongly suppressed **(Figure 4G)**. This result is similar to the response of NOS mutants in cPRC, suggesting that NIT-GC1 mediates depolarising responses after UV stimulation of cPRC by NO and may be involved in extracellular signalling. Interestingly, in NIT-GC2 morphant larvae, an increase in calcium levels occurred immediately after light stimulation **(Figure 4H)**. The different responses of cPRC upon knockdown of NIT-GC1 and NIT-GC2 suggest that there may be different intracellular signalling mechanisms associated with their respective localisations.

@@

@@

therefore compared calcium imaging of postsynaptic cells (INNOS, INRGW, Ser-h1 and MC) in the cPRC circuit during UV stimulation of cPRC between WT and NOS mutant larvae.
